Self-assembly of colloid-cholesteric composites provides a possible route to switchable optical materials

### 2014-04-16 trunk/papers/2012-sandwich/CholColloids.tex
updates following editorial comments plus cover art
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig@2363 40876441-3496-0410-a613-a2fca35b2ae2

\noindent
When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimize the free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures may have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new devices~\cite{colloiddevice,tanakanatmat}.

But one may also choose to disperse colloidal particles in a cholesteric, 
rather than nematic, liquid crystal. The molecules making up this material are
chiral, and this causes their average orientation (described by a director field) to rotate in space in
a helical fashion, rather than remain uniform as in nematics. 
This typically creates a 1-dimensional periodic structure, called the cholesteric phase, whose wavelength 
is the pitch, $p$. The periodicity can however become fully three dimensional in the so-called blue phases (BPs)~\cite{mermin}. BPs arise because the director field can twist around more than one direction at once (see 
the cartoon in Supplementary Fig.~1). Such double twist 
regions (generally cylinders) are energetically favoured at high enough molecular chirality, but
lead to geometric frustration because it is impossible to tile the whole 3D
space with double twist cylinders without also introducing ``disclination lines'' 
(These are singular topological defects on which the director field is 
undefined.) In blue phases these disclination lines themselves either form a 3D periodic regular lattice (in so called BPI and BPII), or remain disordered (BPIII)~\cite{bp3}. 
%
BPs are stabilised by the tendency of chiral molecules to twist, but destabilised by the cost of forming topological defects (which locally diminish the molecular alignment). In conventional cholesteric materials~\cite{mermin} this balance is only achievable in a narrow temperature range of a few degrees, around the onset of liquid crystalline ordering.
\textcolor{red}{
However, recent advances in formulation have widened this stability
range enormously~\cite{kikuchi,bplasers,coleswidetrange,bpdevice}. Most
recently, templated BPs with a stability range of -125
to~125$^\circ$C have been reported \cite{coles}, paving the way to
applications of BPs in operational display devices.}

@@

\textcolor{red}{
The simulations presented here are representative of a liquid crystal
blue phase with unit cell size in the range of 100--500~nm \cite{mermin},
and particles with a diameter of around 50~nm. The simulations
address both the bulk structure, and the structure which forms when
the colloidal dispersion is confined in a narrow sandwich of width
comparable to the unit cell size~\cite{extrareference2}. It is seen
that a wide variety of structure is possible, and may be influenced
by parameters including the solid volume fraction, and the details
of the anchoring at the colloid surface (and the at the surface of the
confining walls). Further, evidence is found for metastable switching of
this structure in an applied electric field. Details of the
computational approach are set out in the Methods section, while
simulation parameters are reported in the Supplementary Information.
}

@@

\noindent
\textbf{Structure in bulk BPI.}
We show in Fig.~1 A-D the final state structures of four simulations in which we first equilibrated a stable BPI disclination network in bulk (with periodic boundary conditions) and then dispersed colloidal particles at random within it. Among dimensionless control parameters are the particle solid volume fraction $\phi$  ($\phi = 1\%$ or $5\%$); the ratio $r = R/\lambda$ of particle radius to the BP lattice parameter $\lambda$ (here $r\simeq 0.14$) and the ratio $w = WR/K$ where $W$ is the anchoring strength of the cholesteric at the particle surface, and $K$ is the elastic constant of the liquid crystal.
\textcolor{red}{
The anchoring strength is defined such that $W > 0$ imparts a preferred
orientation to the director field at the particle surface. This orientation
can be either normal to the surface or in a plane tangential to the surface,
a property related to surface chemistry in experimental systems.% but which
%can be imposed at will in simulations.
We choose} $w = 0.23, 2.3$ which are within typical experimental range~\cite{tiffany}. This parameter can be viewed as the ratio of the anchoring energy, $WR^2$, to the elastic energy scale for distortion~$KR$.

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloidal-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small (Fig.~ 1A,B; Supplementary Movies~1 and~2), the attraction of particles to the disclination lattice causes
 the lattice to become covered by particles, with little disruption of its
 long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloidal particles which leads to very slow dynamics
(Supplementary Fig.~4) and the formation of small colloidal lines within disclinations, especially for dilute samples (Fig. 1A). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

For large $w$ (Fig.~1C,D; Supplementary Movies~3 and~4), the picture changes
dramatically.
Each colloidal particle now disrupts significantly the order of the nearby fluid, with defect formation close to its surface~\cite{stark}. The resulting strong disruption of local liquid crystalline order around each particle leads now to 
a strong interparticle attraction. This restructures the disclination
network completely, destroying the long range order of
the original BPI topology. Most notably, particle aggregation favours
the formation of defect junctions, where four disclinations meet. This structural motif is present in the unit cell of BPII rather than BPI (and also seen in the amorphous BPIII~\cite{bp3}). The particle clusters are disjoint at low volume fractions ($\phi\simeq 1\%$) but they interact elastically via the connecting disclinations. At larger $\phi$, the aggregates join up to form a percolating colloidal cluster which essentially templates the disclinations rather than vice versa.

As colloidal-nematic composites~\cite{tiffany}, all the structures in
Fig.~1A-D 
should be soft solids, of nonzero elastic 
modulus $G$ at low frequency. This contrasts with pure BPs, where
$G=0$ as the fluid can flow with a finite
viscosity via permeation of molecules through a fixed defect lattice~\cite{permeation1,permeation2}. We expect the
structures in Fig.~1A-C to be only weak gels, as here the 
defect network percolates but particle contacts do not. This is broadly similar to the cholesteric-nanoparticle gel reported
in~\cite{lubensky} (with $G<1$ Pa). In contrast, the colloidal gel
structure in Fig.~1D should be much more resistant mechanically, 
due to the formation of a thick percolating particle network. The 
structure is now akin to that of the self-quenched glass found in 
colloidal-nematic composites~\cite{tiffany} at much higher volume 
fraction (over 20\% as opposed to below 5\% here), for which $G\sim 10^3$ Pa.
\textcolor{red}{
This figure is consistent with an estimate obtained by comparing
the mechanical properties of the network of defects stabilised by
colloidal particles with those of rubber~\cite{ramos} for
the relevant simulation parameters (see Supplementary Information).}

A more practical protocol for including particles in a BPI phase is to disperse the particles at random in an isotropic phase, and then quenching this into the temperature range where BPI is stable. Here we know that, without particles, the kinetics can favour disordered (BPIII-like) intermediates which may be long lived or even metastable~\cite{domaingrowth}. Fig.~1E-H shows outcomes of such a quench protocol with the same parameter sets as in Fig.~1A-D. For small $w$, the particles again decorate the disclination lines, but are this time templated onto an amorphous network rather than an ordered one.
Increasing the volume fraction (Fig.~1F) leads to a more even coverage with
a better defined particle spacing. 
For large $w$, on the other hand, strong interparticle forces again dictate
the self-assembly. At low concentrations, discrete particle clusters decorate an amorphous defect lattice (Fig.~1G) whereas at higher $\phi$ a percolating network of colloidal strands (the gel in Fig.~1H) is again seen. 
The structures in Fig. 1A-H have been followed for $\sim$20 ms 
(see Supplementary Fig.~4); they correspond to either very slowly evolving
states or  deep metastable minima (elastic interactions are of
order $KR$ and dwarf thermal motion, see
Supplementary Fig.~4 and Ref.~\cite{stark,lavrentovich}).

@@

Fig.~2A-D shows the case of normal wall anchoring on varying $\phi$ and $w$. We quench from the isotropic to the ordered phase as was done in Fig.~1E-H, and again find a metastable, amorphous disclination network (see Supplementary Information). The resulting colloidal-cholesteric composites are somewhat similar to those seen in the bulk: for small $w$ the particles are templated by the network (Fig.~2A,B), whereas for large $w$ we observe isolated clusters (Fig.~2C) and percolated colloidal gels (Fig.~2D), for $\phi = 1\%$ and $4\%$ respectively. There are, however, significant effects of the confining walls.
First, the normal anchoring at the wall recruits particles there,  giving (for the chosen $h$) a density enhancement of the colloid particles at both walls 
(see side-views in Fig.~2A-D). Second, the clusters formed at large $w$ are themselves anisotropic with an oblate habit. These clusters
are similar to those reported experimentally for large colloid particles in a
cholesteric (but non-BP) host~\cite{niek}.

Fig.~2E-H shows comparable simulations for planar wall anchoring. 
Without colloidal particles, planar anchoring results in a regular
cholesteric helix,
with axis perpendicular to the boundary walls. 
This twisted texture develops via a set of defect loops,
which shrink and disappear to leave a defect-free sample. However,
dispersing particles with small $w$ arrests the loop shrinking process,
stabilising twisted rings (Fig.~2E) at small $\phi$, and more complicated 
double-stranded structures at larger $\phi$ (Fig.~2F).
For larger $w$, we observe thick colloidal ``ropes" about 5 particles across
(Fig.~2G,H). These ropes undulate to follow the chiral structure of the 
underlying fluid. On increasing the density they develop branch points and finally again approach a percolating gel morphology. Note that planar wall anchoring is incompatible with the normal anchoring on the colloidal particles (for $w>0$), which are thus repelled from both walls (see the side views in Fig.~2E-H).

\textcolor{red}{
The degree of similarity between
the bulk and confined structures is somewhat masked by the local
change in particle density in the confined case. If one allows for
this, Figs.~1 and~2 show the individual structural motifs are actually
rather similar: templated lines for weak anchoring and clusters for
strong anchoring.
}

\medskip
\noindent
\textbf{Switching in electric field.}
An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous Blue Phase III,
this time at a higher chirality. Dispersing colloidal particles in this
network
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3A; Supplementary Movie~5). Applying a high enough
field 
along the $x$ direction leads to a switching to a new phase, where colloidal 
arrays fit within a network with distorted hexagonal cells (Fig.~3B 
and Supplementary Fig.~7; Supplementary Movie~6). 
Remarkably, upon field removal this honeycomb structure does not find its 
way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Supplementary Fig.~7), with residual anisotropy along the field
switching 
direction. Cycling the field along the $x$ and $y$ directions leads to
distinct metastable structures, which are retained after field removal
(Supplementary Fig.~8 and Fig.~9).

@@

Here, the simulation method is described. Note that
all quantities are expressed in simulation units. Details
of the simulation parameters, and their mapping to physical units,
may be found in the Supplementary Information.

@@

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_a}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_a$ (here assumed positive) is the dielectric anisotropy of the liquid
crystal. The strength of the electric field is quantified via
one further dimensionless number
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_a}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}
The quantity ${\cal E}$ is known as the reduced field strength.

\medskip
\noindent
\textbf{Surface free energy.}
In addition to the fluid free energy density $f(Q_{\alpha\beta})$, a surface
free energy density is also present to represent the energetic cost of
anchoring at a solid surface. If the preferred order parameter tensor
at the solid surface is $Q^0_{\alpha\beta}$, then the surface free energy
density (per unit area) is
\begin{equation}
f_s(Q_{\alpha\beta}, Q^0_{\alpha\beta})
= {\textstyle \frac{1}{2}}W(Q_{\alpha\beta} - Q^0_{\alpha\beta})^2,
\end{equation}
where $W$ is a constant determining the strength of the anchoring.
The determination of $Q^0_{\alpha\beta}$ for normal and planar anchoring is
discussed below. For a colloidal particle of radius $R$, the strength of
the surface
anchoring compared which the bulk fluid elastic constant may be quantified
by the dimensionless parameter $WR/K$. For small values of this parameter,
the presence of a particle surface should have little impact on the local
fluid LC ordering.

\medskip
\noindent
\textbf{Dynamics of the order parameter.}
A framework for the dynamics of liquid crystals is provided by the 
Beris-Edwards model \cite{beris}, in which the time evolution of the
tensor order parameter obeys
\begin{equation}
\label{eom}
\left(\partial_t+ u_\nu \partial_\nu \right) Q_{\alpha\beta} - S_{\alpha\beta}
= \Gamma H_{\alpha\beta}.
\end{equation}
In the absence of flow, Eq.~\ref{eom} describes a relaxation towards
equilibrium on a timescale determined by a collective rotational diffusion 
constant $\Gamma$. This relaxation is driven by the molecular field
$H_{\alpha\beta}$, which is the functional derivative of the free energy
with respect to the order parameter \cite{beris}:
\begin{equation}
H_{\alpha\beta} = -\frac{\delta {\cal F}} {\delta Q_{\alpha\beta}} 
+ {\textstyle \frac{1}{3}} \delta_{\alpha\beta} 
\mathrm{Tr}\left(\frac{\delta {\cal F}}{\delta Q_{\alpha\beta}}\right).
\end{equation}

The tensor $S_{\alpha\beta}$ in Eq.~\ref{eom} completes the material
derivative for rod-like molecules \cite{beris}. It couples the order
parameter to the symmetric and antisymmetric parts of the velocity 
gradient tensor $W_{\alpha \beta}\equiv\partial_\beta u_\alpha$.
The symmetric part $A_{\alpha\beta}$ and the antisymmetric part
$\Omega_{\alpha\beta}$ are defined as 
\begin{equation}
A_{\alpha\beta} = {\textstyle \frac{1}{2}} (W_{\alpha\beta} + W_{\beta\alpha}),
\end{equation}
and
\begin{equation}
\Omega_{\alpha\beta} = {\textstyle \frac{1}{2}} (W_{\alpha\beta} - W_{\beta\alpha}).
\end{equation}
This full coupling term is then
\begin{eqnarray}
\label{coupling-term}
S_{\alpha\beta}  &=&
(\xi A_{\alpha\nu} + \Omega_{\alpha\nu})
(Q_{\nu\beta} + {\textstyle \frac{1}{3}}\delta_{\nu\beta})
+
(Q_{\alpha\nu} + {\textstyle \frac{1}{3}} \delta_{\alpha\nu})
(\xi A_{\nu\beta} - \Omega_{\nu\beta})
\nonumber\\ 
&-&2 \xi ({Q_{\alpha\beta} + {\textstyle \frac{1}{3}}\delta_{\alpha\beta}})
Q_{\nu\nu} W_{\nu\nu}.
\end{eqnarray}
Here, $\xi$ is a material-dependent that controls the
relative importance of rotational and elongational flow for molecular
alignment, and determines in practice how the orientational order
responds to a local shear flow. 
In all cases, the value used here is $\xi = 0.7$, which is within
the `flow aligning' regime, where molecules align at a fixed angle
(the Leslie angle) to the flow direction in weak simple shear \cite{deGennes}.
The value of the collective rotational diffusion used in all simulations is
$\Gamma = 0.5$ in simulation units.

@@

In all simulations the colloidal (hard sphere) radius is $R = 2.3$. To
prevent particles overlapping at the hard sphere radius, an additional
short range soft potential is included. This takes the form of
$V(h) = \epsilon (\sigma/h)$ where $h$ is the surface to surface separation.
The parameters are $\epsilon = 0.0004$ and $\sigma = 0.1$ in simulation
units. Both potential and resulting force are smoothly matched to zero at
a cut-off distance of $h_c = 0.25$ simulation units.

@@

\bibitem{fukuda}
Fukuda, J.-I., and Zumer, S.,
Quasi-two-dimensional Skyrmion lattices in a chiral nematic liquid crystal,
{\it Nat. Commun.} {\bf 2}, doi:10.1038/ncomms1250 (2011).

>>>>

\noindent
When spherical colloidal particles are mixed into a nematic liquid crystal, 
they disrupt the orientational order in the fluid and create defects (disclination lines) in the nematic close to their surfaces. To minimise the free energy cost, it is generally advantageous for defects to be shared:
therefore the colloidal particles have a generic tendency to aggregate. Such aggregates may further 
self-assemble into lines~\cite{wiresmiha}, 2D crystals~\cite{zumer}, 
planar structures~\cite{tanaka}, or 3D amorphous glasses~\cite{tiffany}.
Besides being of fundamental interest to materials science, these
structures may have tunable elastic and optical properties~\cite{stark}. Hence 
they offer exciting prospects for applications as biosensors~\cite{abbott}, or
as new devices~\cite{colloiddevice,tanakanatmat}.

But one may also choose to disperse colloidal particles in a cholesteric, 
rather than nematic, liquid crystal. The molecules making up this material are
chiral, and this causes their average orientation (described by a director field) to rotate in space in
a helical fashion, rather than remain uniform as in nematics. 
This typically creates a 1-dimensional periodic structure, called the cholesteric phase, whose wavelength 
is the pitch, $p$. The periodicity can however become fully three dimensional in the so-called blue phases (BPs)~\cite{mermin}. BPs arise because the director field can twist around more than one direction at once (see 
the cartoon in Supplementary Fig.~1). Such double twist 
regions (generally cylinders) are energetically favoured at high enough molecular chirality, but
lead to geometric frustration because it is impossible to tile the whole 3D
space with double twist cylinders without also introducing ``disclination lines'' 
(These are singular topological defects on which the director field is 
undefined.) In blue phases these disclination lines themselves either form a 3D periodic regular lattice (in so called BPI and BPII), or remain disordered (BPIII)~\cite{bp3}. 
%
BPs are stabilised by the tendency of chiral molecules to twist, but destabilised by the cost of forming topological defects (which locally diminish the molecular alignment). In conventional cholesteric materials~\cite{mermin} this balance is only achievable in a narrow temperature range of a few degrees, around the onset of liquid crystalline ordering.
However, recent advances in formulation have widened this stability
range enormously~\cite{kikuchi,bplasers,coleswidetrange,bpdevice}. Most
recently, templated BPs with a stability range of -125
to~125$^\circ$C have been reported \cite{coles}, paving the way to
applications of BPs in operational display devices.

@@

The simulations presented here are representative of a liquid crystal
blue phase with unit cell size in the range of 100--500~nm \cite{mermin},
and particles with a diameter of around 50~nm. The simulations
address both the bulk structure, and the structure which forms when
the colloidal dispersion is confined in a narrow sandwich of width
comparable to the unit cell size~\cite{extrareference2}. It is seen
that a wide variety of structure is possible, and may be influenced
by parameters including the solid volume fraction, and the details
of the anchoring at the colloid surface (and the at the surface of the
confining walls). Further, evidence is found for metastable switching of
this structure in an applied electric field. Details of the
computational approach are set out in the Methods section, while
simulation parameters are reported in the Supplementary Methods.

@@

\noindent
\textbf{Structure in bulk BPI.}
We show in Fig.~1~(\textbf{a}--\textbf{d})
the final state structures of four simulations in which we first equilibrated a stable BPI disclination network in bulk (with periodic boundary conditions) and then dispersed colloidal particles at random within it. Among dimensionless control parameters are the particle solid volume fraction $\phi$  ($\phi = 1\%$ or $5\%$); the ratio $r = R/\lambda$ of particle radius to the BP lattice parameter $\lambda$ (here $r\simeq 0.14$) and the ratio $w = WR/K$ where $W$ is the anchoring strength of the cholesteric at the particle surface, and $K$ is the elastic constant of the liquid crystal.
The anchoring strength is defined such that $W > 0$ imparts a preferred
orientation to the director field at the particle surface. This orientation
can be either normal to the surface or in a plane tangential to the surface,
a property related to surface chemistry in experimental systems.
We choose $w = 0.23, 2.3$ which are within typical experimental range~\cite{tiffany}. This parameter can be viewed as the ratio of the anchoring energy, $WR^2$, to the elastic energy scale for distortion~$KR$.

For a single colloidal particle in a nematic fluid, $w$ controls the formation of a local topological defect at the particle surface~\cite{stark}, with weak distortion of the nearby fluid at small $w$. In the colloidal-BP system, $w$ also plays a determining role in the final composite structure. When $w$ is small
(Fig.~1\textbf{a} and Fig.~1\textbf{b}; Supplementary Movies~1 and~2),
the attraction of particles to the disclination lattice causes
 the lattice to become covered by particles, with little disruption of its
 long-range order. Thus a templated colloidal crystal is created, as first reported in~\cite{miha}. Interestingly, the weak but finite anchoring apparently creates an elastic force between the colloidal particles which leads to very slow dynamics
(Supplementary Fig.~4) and the formation of small colloidal lines within disclinations, especially for dilute samples (Fig.~1\textbf{a}). It seems likely that polymer-stabilised  blue phases~\cite{kikuchi} work by a similar principle to this small $w$ case, with a weak segregation of monomers towards the defect regions.

For large $w$ (Fig.~1\textbf{c} and Fig.~1\textbf{d};
Supplementary Movies~3 and~4), the picture changes
dramatically.
Each colloidal particle now disrupts significantly the order of the nearby fluid, with defect formation close to its surface~\cite{stark}. The resulting strong disruption of local liquid crystalline order around each particle leads now to 
a strong interparticle attraction. This restructures the disclination
network completely, destroying the long range order of
the original BPI topology. Most notably, particle aggregation favours
the formation of defect junctions, where four disclinations meet. This structural motif is present in the unit cell of BPII rather than BPI (and also seen in the amorphous BPIII~\cite{bp3}). The particle clusters are disjoint at low volume fractions ($\phi\simeq 1\%$) but they interact elastically via the connecting disclinations. At larger $\phi$, the aggregates join up to form a percolating colloidal cluster which essentially templates the disclinations rather than vice versa.

As colloidal-nematic composites~\cite{tiffany}, all the structures in
Fig.~1~(\textbf{a}--\textbf{d}) 
should be soft solids, of nonzero elastic 
modulus $G$ at low frequency. This contrasts with pure BPs, where
$G=0$ as the fluid can flow with a finite
viscosity via permeation of molecules through a fixed defect lattice~\cite{permeation1,permeation2}. We expect the
structures in Fig.~1~(\textbf{a}--\textbf{c}) to be only weak gels,
as here the 
defect network percolates but particle contacts do not. This is broadly similar to the cholesteric-nanoparticle gel reported
in~\cite{lubensky} (with $G<1$ Pa). In contrast, the colloidal gel
structure in Fig.~1\textbf{d} should be much more resistant mechanically, 
due to the formation of a thick percolating particle network. The 
structure is now akin to that of the self-quenched glass found in 
colloidal-nematic composites~\cite{tiffany} at much higher volume 
fraction (over 20\% as opposed to below 5\% here), for which $G\sim 10^3$ Pa.
This figure is consistent with an estimate obtained by comparing
the mechanical properties of the network of defects stabilised by
colloidal particles with those of rubber~\cite{ramos} for
the relevant simulation parameters (see Supplementary Table~1).

A more practical protocol for including particles in a BPI phase is to
disperse the particles at random in an isotropic phase, and then
quenching this into the temperature range where BPI is stable.
Here we know that, without particles, the kinetics can favour
disordered (BPIII-like) intermediates which may be long lived
or even metastable~\cite{domaingrowth}. Fig.~1~(\textbf{e}--\textbf{h})
shows outcomes of such a quench protocol with the same parameter sets as
in Fig.~1~(\textbf{a}--\textbf{d}). For small $w$, the particles again
decorate the disclination lines, but are this time templated onto an
amorphous network rather than an ordered one.
Increasing the volume fraction (Fig.~1\textbf{f}) leads to a more even
coverage with a better defined particle spacing. 
For large $w$, on the other hand, strong interparticle forces again dictate
the self-assembly. At low concentrations, discrete particle clusters
decorate an amorphous defect lattice (Fig.~1\textbf{g}) whereas at
higher $\phi$ a percolating network of colloidal strands
(the gel in Fig.~1\textbf{h}) is again seen. 
The structures in Fig.~1~(\textbf{a}--\textbf{h}) have been followed for
$\sim$20 ms 
(see Supplementary Fig.~4); they correspond to either very slowly evolving
states or  deep metastable minima (elastic interactions are of
order $KR$ and dwarf thermal motion, see
Supplementary Fig.~4 and Ref.~\cite{stark,lavrentovich}).

It may be possible to anneal the structures obtained via the quench route,
although we have not attempted this in simulation. The results presented
suggest that even when starting with colloidal particles randomly dispersed
in an ordered blue phase, the particles are able to distort the structure
(Fig.~1~\textbf{a},\textbf{b}). This suggests that it would be difficult
to identify a quench
protocol which would lead to less, rather than more, disorder.

@@

Fig.~2~(\textbf{a}--\textbf{d}) shows the case of normal wall anchoring on
varying $\phi$ and $w$. We quench from the isotropic to the ordered phase
as was done in Fig.~1~(\textbf{e}--\textbf{h}), and again find a metastable,
amorphous disclination network
(see Supplementary Methods). The resulting colloidal-cholesteric composites
are somewhat similar to those seen in the bulk: for small $w$ the particles
are templated by the network (Fig.~2\textbf{a},\textbf{b}), whereas for
large $w$ we observe isolated clusters (Fig.~2\textbf{c}) and percolated
colloidal gels (Fig.~2\textbf{d}), for $\phi = 1\%$ and $4\%$ respectively.
There are, however, significant effects of the confining walls.
First, the normal anchoring at the wall recruits particles there,  giving
(for the chosen $h$) a density enhancement of the colloid particles at
both walls  (see side-views in Fig.~2\textbf{a}--\textbf{d}). Second,
the clusters formed at large $w$ are themselves anisotropic with an
oblate habit. These clusters
are similar to those reported experimentally for large colloid particles in a
cholesteric (but non-BP) host~\cite{niek}.

Fig.~2~(\textbf{e}--\textbf{h}) shows comparable simulations for planar wall
anchoring.  Without colloidal particles, planar anchoring results in a
regular cholesteric helix,
with axis perpendicular to the boundary walls. 
This twisted texture develops via a set of defect loops,
which shrink and disappear to leave a defect-free sample. However,
dispersing particles with small $w$ arrests the loop shrinking process,
stabilising twisted rings (Fig.~2\textbf{e}) at small $\phi$, and more
complicated  double-stranded structures at larger $\phi$ (Fig.~2\textbf{f}).
For larger $w$, we observe thick colloidal ``ropes" about 5 particles across
(Fig.~2\textbf{g},\textbf{h}). These ropes undulate to follow the chiral
structure of the 
underlying fluid. On increasing the density they develop branch points and
finally again approach a percolating gel morphology. Note that planar wall
anchoring is incompatible with the normal anchoring on the colloidal
particles (for $w>0$), which are thus repelled from both walls
(see the side views in Fig.~2\textbf{e}--\textbf{h}).

The degree of similarity between
the bulk and confined structures is somewhat masked by the local
change in particle density in the confined case. If one allows for
this, Figs.~1 and~2 show the individual structural motifs are actually
rather similar: templated lines for weak anchoring and clusters for
strong anchoring.

\medskip
\noindent
\textbf{Switching in electric field.}
An interesting question, relevant for applications to devices, is whether 
these structures can be modified, or switched, with electric fields.
Fig.~3 addresses this question, starting from amorphous blue phase~III,
this time at a higher chirality. Dispersing colloidal particles in this
network
results in the particle occupying the available nodes, where the order
parameter is low (Fig.~3\textbf{a}; Supplementary Movie~5).
Applying a high enough field 
along the $x$ direction leads to a switching to a new phase, where colloidal 
arrays fit within a network with distorted hexagonal cells (Fig.~3\textbf{b} 
and Supplementary Fig.~7; Supplementary Movie~6). 
Remarkably, upon field removal this honeycomb structure does not find its 
way back to the 
configuration before switching on the field, but gets stuck into a metastable 
structure (Supplementary Fig.~7), with residual anisotropy along the field
switching 
direction. Cycling the field along the $x$ and $y$ directions leads to
distinct metastable structures, which are retained after field removal
(Supplementary Fig.~8 and Fig.~9).
We note that for the fluid parameters used here, the field induced
changes are irreversible with the weakly anchoring colloid particles.
The mesophase without colloid particles undergoes
a reversible ordering transition~\cite{bp3}; the reversibility
is retained in the
presence of particles only if their anchoring is very weak.

@@

Here, the simulation method is described. Note that
all quantities are expressed in simulation units. Details
of the simulation parameters, and their mapping to physical units,
may be found in Supplementary Table~1 and the Supplementary Methods.

@@

The effect of an electric field can be modelled by including the following
term in the bulk free energy density:
\begin{equation}
-\frac{\epsilon_\mathrm{a}}{12\pi} E_{\alpha}Q_{\alpha\beta}E_{\beta},
\end{equation} 
where $E_{\alpha}$ are the components of the electric field, and 
$\epsilon_\mathrm{a}$ (here assumed positive) is the dielectric anisotropy
of the liquid
crystal. The strength of the electric field is quantified via
one further dimensionless number
\begin{equation}
{\cal E}^2 = \frac{27 \epsilon_\mathrm{a}}{32 \pi A_0 \gamma} E_\alpha E_\alpha.
\end{equation}
The quantity ${\cal E}$ is known as the reduced field strength.

\medskip
\noindent
\textbf{Surface free energy.}
In addition to the fluid free energy density $f(Q_{\alpha\beta})$, a surface
free energy density is also present to represent the energetic cost of
anchoring at a solid surface.
In the case of normal anchoring, the surface free energy
density (per unit area) is
\begin{equation}
f_s(Q_{\alpha\beta}, Q^0_{\alpha\beta})
= {\textstyle \frac{1}{2}}W(Q_{\alpha\beta} - Q^0_{\alpha\beta})^2,
\end{equation}
where $Q^0_{\alpha\beta}$ is the preferred order
parameter tensor at the solid surface, and $W$ is a constant determining
the strength of the anchoring. In the case of planar anchoring, a slightly
more complicated expression is required to allow for degeneracy
(see, e.g., \cite{fournier2005}).
The determination of $Q^0_{\alpha\beta}$ for normal and planar anchoring is
discussed below. For a colloidal particle of radius $R$, the strength of
the surface
anchoring compared which the bulk fluid elastic constant may be quantified
by the dimensionless parameter $WR/K$. For small values of this parameter,
the presence of a particle surface should have little impact on the local
fluid LC ordering.

\medskip
\noindent
\textbf{Dynamics of the order parameter.}
A framework for the dynamics of liquid crystals is provided by the 
Beris-Edwards model \cite{beris}, in which the time evolution of the
tensor order parameter obeys
\begin{equation}
\label{eom}
\left(\partial_t+ u_\nu \partial_\nu \right) Q_{\alpha\beta} - S_{\alpha\beta}
= \mathit{\Gamma} H_{\alpha\beta}.
\end{equation}
In the absence of flow, Eq.~\ref{eom} describes a relaxation towards
equilibrium on a timescale determined by a collective rotational diffusion 
constant $\mathit{\Gamma}$. This relaxation is driven by the molecular field
$H_{\alpha\beta}$, which is the functional derivative of the free energy
with respect to the order parameter \cite{beris}:
\begin{equation}
H_{\alpha\beta} = -\frac{\delta {\cal F}} {\delta Q_{\alpha\beta}} 
+ {\textstyle \frac{1}{3}} \delta_{\alpha\beta} 
\mathrm{Tr}\left(\frac{\delta {\cal F}}{\delta Q_{\alpha\beta}}\right).
\end{equation}

The tensor $S_{\alpha\beta}$ in Eq.~\ref{eom} completes the material
derivative for rod-like molecules \cite{beris}. It couples the order
parameter to the symmetric and antisymmetric parts of the velocity 
gradient tensor $W_{\alpha \beta}\equiv\partial_\beta u_\alpha$.
The symmetric part $A_{\alpha\beta}$ and the antisymmetric part
$\Omega_{\alpha\beta}$ are defined as 
\begin{equation}
A_{\alpha\beta} = {\textstyle \frac{1}{2}} (W_{\alpha\beta} + W_{\beta\alpha}),
\end{equation}
and
\begin{equation}
\Omega_{\alpha\beta} = {\textstyle \frac{1}{2}} (W_{\alpha\beta} - W_{\beta\alpha}).
\end{equation}
This full coupling term is then
\begin{eqnarray}
\label{coupling-term}
S_{\alpha\beta}  &=&
(\xi A_{\alpha\nu} + \Omega_{\alpha\nu})
(Q_{\nu\beta} + {\textstyle \frac{1}{3}}\delta_{\nu\beta})
+
(Q_{\alpha\nu} + {\textstyle \frac{1}{3}} \delta_{\alpha\nu})
(\xi A_{\nu\beta} - \Omega_{\nu\beta})
\nonumber\\ 
&-&2 \xi ({Q_{\alpha\beta} + {\textstyle \frac{1}{3}}\delta_{\alpha\beta}})
Q_{\nu\nu} W_{\nu\nu}.
\end{eqnarray}
Here, $\xi$ is a material-dependent that controls the
relative importance of rotational and elongational flow for molecular
alignment, and determines in practice how the orientational order
responds to a local shear flow. 
In all cases, the value used here is $\xi = 0.7$, which is within
the `flow aligning' regime, where molecules align at a fixed angle
(the Leslie angle) to the flow direction in weak simple shear \cite{deGennes}.
The value of the collective rotational diffusion used in all simulations is
$\mathit{\Gamma} = 0.5$ in simulation units.

@@

In all simulations the colloidal (hard sphere) radius is $R = 2.3$. To
prevent particles overlapping at the hard sphere radius, an additional
short range soft potential is included. This takes the form of
$V(h) = \epsilon (\sigma/h)$ where $h$ is the surface to surface separation.
The parameters are $\epsilon = 0.0004$ and $\sigma = 0.1$ in simulation
units. Both potential and resulting force are smoothly matched to zero at
a cut-off distance of $h_\mathrm{c} = 0.25$ simulation units.

@@

\bibitem{fukuda}
Fukuda, J.-I., and Zumer, S.,
Quasi-two-dimensional Skyrmion lattices in a chiral nematic liquid crystal,
{\it Nat. Commun.} {\bf 2}, 246 (2011).
